
Most of the research presented in this thesis has been done in collaboration with other colleagues,
as detailed in the following.

The results presented in chapters~\ref{ch:nnpdf_methodology} and \ref{ch:th_error}
have been produced within the NNPDF collaboration. The original publications are

NNPDF Collaboration, R.~Abdul~Khalek et~al., {\it {A first determination
  of parton distributions with theoretical uncertainties}},  {\em Eur. Phys.
  J.} {\bf C} (2019) 79:838, [\href{http://arxiv.org/abs/1905.04311}{{\tt
  arXiv:1905.04311}}]

NNPDF Collaboration, R.~Abdul~Khalek et~al., {\it {Parton Distributions
  with Theory Uncertainties: General Formalism and First Phenomenological
  Studies}},  {\em Eur. Phys. J. C} {\bf 79} (2019), no.~11 931,
  [\href{http://arxiv.org/abs/1906.10698}{{\tt arXiv:1906.10698}}]\,,

in collaboration with Rabah Abdul Khalek, Richard D. Ball, Stefano Carrazza, Stefano Forte,
Zahari Kassabov, Emanuele R. Nocera, Rosalyn L. Pearson, Juan Rojo, Luca Rottoli,
Maria Ubiali, Cameron Voisey and Michael Wilson.


Results reported in chapter~\ref{ch:jets} were first presented in 

R.~Abdul~Khalek et~al., {\it {Phenomenology of NNLO jet production at the LHC
  and its impact on parton distributions}},  {\em Eur. Phys. J. C} {\bf 80}
  (2020), no.~8 797, [\href{http://arxiv.org/abs/2005.11327}{{\tt
  arXiv:2005.11327}}]\,,

in collaboration with Rabah Abdul Khalek, Stefano Forte, Thomas Gehrmann, Aude Gehrmann-De Ridder,
Nigel Glover, Alexander Huss, Emanuele R. Nocera, Joao Pires, Juan Rojo and Giovanni Stagnitto.

Chapter~\ref{ch:bottom} is based on

S.~Forte, T.~Giani, and D.~Napoletano, {\it {Fitting the b-quark PDF as a
  massive-b scheme: Higgs production in bottom fusion}},  {\em Eur. Phys. J. C}
  {\bf 79} (2019), no.~7 609, [\href{http://arxiv.org/abs/1905.02207}{{\tt
  arXiv:1905.02207}}]

in collaboration with Stefano Forte and Davide Napoletano.

Chapter~\ref{ch:scalar_model} is based on

L.~Del~Debbio, T.~Giani, and C.~J. Monahan, {\it {Notes on lattice observables
  for parton distributions: nongauge theories}},  {\em JHEP} {\bf 09} (2020)
  021, [\href{http://arxiv.org/abs/2007.02131}{{\tt arXiv:2007.02131}}]\,,

in collaboration with Luigi Del Debbio and Christopher J. Monahan.

Chapter~\ref{ch:qpdfNNPDF} is based on

K.~Cichy, L.~Del~Debbio, and T.~Giani, {\it {Parton distributions from lattice
  data: the nonsinglet case}},  {\em JHEP} {\bf 10} (2019) 137,
  [\href{http://arxiv.org/abs/1907.06037}{{\tt arXiv:1907.06037}}]\,,

in collaboration with Krzysztof Cichy and Luigi Del Debbio.

Finally Chapter~\ref{ch:ppdf_NNPDF} is based on

L.~Del~Debbio, T.~Giani, J.~Karpie, K.~Orginos, A.~Radyushkin, and
  S.~Zafeiropoulos, {\it {Neural-network analysis of Parton Distribution
  Functions from Ioffe-time pseudodistributions}},  {\em JHEP} {\bf 02} (2021)
  138, [\href{http://arxiv.org/abs/2010.03996}{{\tt arXiv:2010.03996}}]

in collaboration with Luigi Del Debbio, Joseph Karpie, Kostas Orginos, Anatoly
Radyushkinc and Savvas Zafeiropoulos.

When results are presented I tried to report more details where I believe that I have played a major role,
reporting additional material when it is necessary for the general understanding of the discussion. 


%
In order to write the first two chapters, devoted to a general overview of some basic concepts of QCD and factorization,
along with the cited references I have also referred to a number of lectures I have attended during the first year of my PhD
at the University of Edinburgh.
In particular, I want to mention the 2017/2018 QCD course held by Prof. Einan Gardi 
and the 2017/2018 lattice QCD lectures given by Prof. Antonin Portelli. 

Unless specified in the caption, I have produced all the figures reported in this thesis.


