In the following I report a draft of the table of contents of my thesis.
For most of the sections I have already collected all the material I need, expect for some of the points of the third chapter,
which are still work in progress. However I have not started the proper writing of the thesis yet.

\begin{enumerate}
    \item Factorization theorems in QCD
    \begin{itemize}
        \item Basics of QCD: I will revise some basic notions of QCD, from the lagrangian to the running coupling.
        \item Factorization theorems in high-energy physics: I will revise and discuss the parton model and factorization theorems
        in high-energy physics, focusing on DIS and hadronic processes.
        \item Factorization for heavy quarks: I will discuss massive and massless schemes and the FONLL method in the computation of
        physical cross sections.
    \end{itemize}
    \item Towards NNPDF4.0
    \begin{itemize}
        \item Machine learning in physics: I will discuss applications of machine learning in physics,
        introducing a series of general concepts and techniques commonly used in AI literature.
        In order to do this I will present part of the work done in Ref.~\cite{Cossu:2018pxj}, where such
        tools have been applied to a statistical physics problem.
        I will then move to discuss how machine learning has been applied within the framework
        of high-energy physics, to determine PDFs within the NNPDF collaboration.
        \item N3fit development: I will present studies regarding the choice of basis and parameterization in a PDFs fit.
        I will also discuss the phenomenological impact of positivity and integrability constraints on PDFs.
        \item Theory uncertainty: I will discuss the treatment of theory uncertainty in a fit, with particular focus on the details 
        of its practical implementation. Based on \cite{AbdulKhalek:2019bux, AbdulKhalek:2019ihb}.
        \item New experimental data: I will present and discuss the impact of new experimental data within the standard NNPDF31 dataset,
        focusing on jets data. Based on \cite{AbdulKhalek:2020jut}.     
        \item A new treatment for the b-PDF: I will present an alternative treatment of the b-PDF, based on \cite{Forte:2019hjc}.
    \end{itemize}
    \item PDFs from lattice data
    \begin{itemize}
        \item Factorization for equal-time correlators: 
        I will discuss the basic theoretical ideas behind the extraction of PDFs from lattice QCD data, presenting
        the main points within the framework of a nongauge theory.
        A paper about this is in preparation.
        \item PDFs from q-PDFs data: I will discuss results regarding PDFs from quasi-PDFs lattice data. Based on \cite{Cichy2019}.
        \item PDFs from p-PDFs data: I will discuss results regarding PDFs from pseudo-PDFs lattice data. 
        A paper about this is in preparation.
        \item Towards a first global lattice QCD fit: I will discuss the possibility of performing a first global analysis 
        over lattice QCD data.
    \end{itemize}
\end{enumerate}